Chess EDA

### 2023-03-15 owen/01_eda.qmd
some initial stuff getting win rate

library(tidyverse)

@@

- Confirm the time controls for the Lichess data?

What types of data?
How many games involve titles?

@@

@@

Get all games from a specific player in a specific format,
to see if `RatingDiff` tracks how
the rating changes.

@@

@@

## have a few different players with good number of games
## get list of all players, fit some sort of model using all players
## who play enough games

head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)

@@

We will identify focal players as those who were used to query the API,
i.e the `Username` for each game.

```{r, past-n-results}

@@

curr_games <-

>>>>

library(tidyverse)
library(RcppRoll)

theme_set(theme_bw())

@@

We first wish to investigate the properties of the scraped data.

@@

We see that the vast majority of games in this dataset 
are Standard, with the regular rules, etc. Similarly, the most
commonly played time controls is 180+0,
which corresponds to 3 minute blitz games with no time increment.

Of standard games, we see that the majority are rated blitz and bullet games.
These correspond to randomly playing available players on the site 
at a time. This differs from tournaments, which make up a sizeable proportion
of standard games also.

@@

Get all games from a specific player in a specific format,
to see that `RatingDiff` tracks how
the rating changes. For the White/Black player based
on the result of the current game.

@@

For all current analysis we will restrict to 1 minute rated bullet games,
the most commonly played format in this dataset, with 13k games compared
to 12k games of 30 second bullet and 12k games of 3 minute blitz.

@@

## have a few different players with good number of games
## get list of all players, fit some sort of model using all players
## who play enough games
```

The top 25 players all play at least 80 games in this format, and across 
all games in this format we see relatively few draws. This is unsurprising
as it it hard to end at a draw in these short time controls, when one player
can lose on time.

```{r}
head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)

@@

We will identify focal players as those who were used to query the API,
i.e the `Username` for each game. We only consider Rated 60 second bullet games
here.

```{r, past-n-results}
#| message: false

@@

curr_user <- "penguingim1"
curr_n <- 10

curr_games <- bullet_60 %>% 
  filter(White == curr_user | Black == curr_user)

hist_games <- curr_games %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
  select(White:BlackElo, focal_white)

## need to check all the game types are the same here then
table(curr_games$Variant)
table(curr_games$Result)

## then get the result for curr_user for each game, there will be draws here

user_games <- hist_games %>% 
  mutate(focal_result = case_when(
    (focal_white == 1 & Result == "1-0") ~ 1,
    (focal_white == 0 & Result == "0-1") ~ 1,
    (Result == "1/2-1/2") ~ 0.5,
    .default = 0
    )) %>% 
  mutate(focal_win_prop = c(rep(0, curr_n - 1), roll_mean(focal_result, n = curr_n)))

user_games %>% 
  ggplot(aes(x = focal_win_prop)) +
  geom_histogram(bins = 20) +
  labs(title = "Win Ratio for Previous 10 Games, one Player")

### now just need to get the rolling average of focal result
### need to pad this output

### could we do this using a map across the players of interest,
### get their games and then put it all back together afterwards?
### think that should work, just need to make it into a function then

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(rep(0, prev_n - 1), 
                              roll_mean(focal_result, n = prev_n)))

  hist_games
}

head(get_hist(user = curr_user, games = bullet_60, prev_n = 10))

## this seems to be working, then do it across a vector of
## players

```{r}
a <- head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)

top_players <- names(a)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5)

tidy_games %>% 
  mutate(focal_player = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_player) %>% 
  count() %>% 
  arrange(-n)

## why does this have 17 players rather than 25 then?
## why are some missing?
